### 2019-03-26 5_evaluation.tex
Apply Alex's comments to section 2.
Apply Alex's comments to section 5.

We now present a comprehensive evaluation of Task Bench on up to 256
Haswell nodes of the Cori supercomputer~\cite{Cori}, a Cray XC40
machine. Cori Haswell nodes have 2 sockets with Intel Xeon E5-2698 v3
processors (total of 32 physical cores per node), 128 GB RAM, and a
Cray Aries interconnect. We use GCC 7.3.0 for compiling all Task Bench
implementations, and (where applicable) the system default MPI
implementation, Cray MPICH 7.7.3. Additional versions and flags for the
various systems are shown in Table~\ref{tab:flags}.

\subsection{Peak Performance and Efficiency}

@@

The vast majority of systems achieve at or near peak FLOPS. Some
systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOPS compared
to systems which share all cores between the application and runtime. Some of the
slower systems struggle to achieve peak FLOPS, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes it prohibitively
expensive to run larger problem sizes. For example, the Spark job in
this configuration ran for over 6 hours.

Figure~\ref{fig:efficiency} shows efficiency (as a percentage of
peak FLOPS achieved) plotted against task granularity. This plot is
used to calculate METG, as described in Section~\ref{sec:metg}. The
red, dashed line shows the point where 50\% efficiency is achieved.

In most cases, task granularity has plateaued prior to the 50\% efficiency point,
though a number of systems continue to improve as efficiency continues
to drop. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold compared to
measuring task scheduling throughput in the presence of empty tasks
(effectively the equivalent of METG(0\%)). Empty tasks permit runtime
systems to make unrealistic tradeoffs which are not applicable to real
application workloads where some degree of efficiency is necessarily a
goal. Reporting numbers with empty tasks therefore gives unfair
advantage to systems that make such unrealistic tradeoffs.

@@

\begin{itemize}
\item Figure~\ref{fig:metg-compute-stencil} is a 1D stencil where each
  task depends on 3 other tasks (including the same point in the
  previous timestep).
\item Figure~\ref{fig:metg-compute-nearest} is a pattern where each
  task depends 5 others, chosen to be as close as possible.
\item Figure~\ref{fig:metg-compute-spread} is a pattern where each
  task depends on 5 others, spread as widely as possible.
\item Figure~\ref{fig:metg-compute-4x-nearest} shows 4 identical
  copies of the nearest dependence pattern executing concurrently.
\end{itemize}

@@

In general, the least complicated pattern (stencil) is most favorable
to MPI, as it represents a truly bulk synchronous task graph with no
opportunity for task parallelism. For the stencil pattern, the
dominating factor is the basic overhead of executing a task, which is
minimal for MPI as the Task Bench implementation is bulk synchronous
and simply executes tasks one after another in alternation with
communication phases. In contrast the asynchronous execution
mechanisms of the other systems is pure overhead in this scenario.

The gap between MPI and other systems shrinks as the complexity of the
communication pattern grows, can even reverse as task parallelism is
added in the form of multiple task graphs.

@@

Most systems for high-performance computing are horizontally scalable,
but this is not true of all the systems included in this
evaluation. Lower is better in Figure~\ref{fig:metg-compute}, and a
flat line is ideal. Systems with non-scalable implementations shows
lines that go up with node count. Most notably, Spark is typically
intended for use in industrial data center applications with task
granularities measured in tens of seconds. Spark uses a centralized
controller which limits throughput, and this is visible in the figure
as line for Spark immediately rises with node count. Keep in mind also
that Spark is being evaluated here with a non-trivial dependence
pattern which is relatively unrepresentative of Spark's normal use
cases. Spark is more efficient with trivial parallelism, as described
in Section~\ref{subsec:number-of-dependencies}.

@@

One advantage of systems with asynchronous execution capabilities is
the ability to mitigate load imbalance, especially in the presence of
task parallelism. To quantify this effect,
Figure~\ref{fig:efficiency-imbalance} plots task granularity vs
efficiency curves under varying amounts of load imbalance. The
imbalanced workloads are generated by randomly varying the task
durations according by a factor determined by the formula $1+(X-0.5)I$
where $X$ is a uniform random variable in the range $[0,1]$ and $I$ is
the degree of imbalance in the range $[0,2]$. The random numbers are
generated with a deterministic pseudo random number generator with a
deterministic seed to ensure that the task durations under all systems
are identical.

The MPI implementation of Task Bench, being bulk synchronous,
obviously suffers the most under load imbalance. The biggest
difference is at large task granularities, where the imbalance has the
effect of putting an upper bound on MPI efficiency. At smaller task
granularities the effect shrinks and may even reverse as systems hit
their fundamental limits due to overhead.

The difference between the remaining systems is due primarily to the
scheduling behavior of the various systems. For example, OpenMP
executes tasks via a thread pool, resulting in a form of work
stealing. This on-node work stealing, in combination with the task
parallelism due to the presence of 4 task graphs, is sufficient to
completely mitigate load imbalance. Systems which exploit task
parallelism but not work stealing on a node achieve lower peak
efficiency, but may achieve smaller task granularities as a result of
not relying on a shared global queue for tasks.

@@

Also of interest is the ability of asynchronous systems to be able to
hide communication latencies in the presence of task
parallelism. Figure~\ref{fig:efficiency-communication} plots task
granularity vs efficiency curves for varying amounts of
communication. The degree of communication is configured by the number
of bytes produced by each task (and therefore communicated with each
task dependency).

>>>>

We present a comprehensive evaluation of Task Bench on up to 256
Haswell nodes of the Cori supercomputer~\cite{Cori}, a Cray XC40
machine. Cori Haswell nodes have 2 sockets with Intel Xeon E5-2698 v3
processors (a total of 32 physical cores per node), 128 GB RAM, and a
Cray Aries interconnect. We use GCC 7.3.0 for compiling all Task Bench
implementations, and (where applicable) the system default MPI
implementation, Cray MPICH 7.7.3. Additional versions and flags for the
various systems are shown in Table~\ref{tab:flags}.

\subsection{Peak Performance and Efficiency}
\label{subsec:peak-performance-and-efficiency}

@@

The vast majority of systems achieve or nearly achieve peak FLOPS. Some
systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOPS compared
to systems which share all cores between the application and runtime. Some of the
slower systems struggle to achieve peak FLOPS, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes it prohibitively
expensive to run larger problem sizes. For example, the Spark job in
this configuration ran for over 6 hours.

Figure~\ref{fig:efficiency} shows efficiency (as a percentage of
peak FLOPS achieved) plotted against task granularity. This plot is
used to calculate METG(50\%), as described in Section~\ref{sec:metg}. The
red, dashed line shows the point where 50\% efficiency is achieved.

In most cases, task granularity has plateaued prior to the 50\% efficiency point,
though some systems continue to improve as efficiency continues
to drop. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold compared to
measuring task scheduling throughput in the presence of empty tasks
(effectively METG(0\%)). Empty tasks permit runtime
systems to make unrealistic tradeoffs which are not applicable to real
application workloads where some degree of efficiency is necessarily a
goal. Reporting numbers with empty tasks therefore gives unfair
advantage to systems that make such unrealistic tradeoffs.

@@

\begin{itemize}
\item Figure~\ref{fig:metg-compute-stencil} is a 1D stencil where each
  task depends on 3 other tasks (including the same point in the
  previous timestep).
\item Figure~\ref{fig:metg-compute-nearest} is a pattern where each
  task depends on 5 others, chosen to be as close as possible.
\item Figure~\ref{fig:metg-compute-spread} is a pattern where each
  task depends on 5 others, spread as widely as possible.
\item Figure~\ref{fig:metg-compute-4x-nearest} shows 4 identical
  copies of the nearest dependence pattern executing concurrently.
\end{itemize}

@@

In general, the least complicated pattern (stencil) is most favorable
to MPI, as it represents a truly bulk synchronous task graph with no
opportunity for task parallelism. For the stencil pattern, the
dominating factor is the basic overhead of executing a task, which is
minimal for MPI as the Task Bench implementation is bulk synchronous
and simply executes tasks one after another in alternation with
communication phases. In contrast the asynchronous execution
mechanisms of the other systems are pure overhead in this scenario.

The gap between MPI and other systems shrinks as the complexity of the
communication pattern grows, and even reverses as task parallelism is
added in the form of multiple task graphs.

@@

Most systems for high-performance computing are horizontally scalable,
but this is not true of all the systems included in this
evaluation. Lower is better in Figure~\ref{fig:metg-compute}, and a
flat line is ideal. Lines that rise with node count indicate less than
ideal scaling. Most notably, Spark is typically
intended for use in industrial data center applications with task
granularities measured in tens of seconds. Spark uses a centralized
controller, which limits throughput, and this is visible in the figure
as line for Spark immediately rises with node count. Keep in mind also
that Spark is being evaluated here with a non-trivial dependence
pattern which is relatively unrepresentative of Spark's normal use
cases. Spark is more efficient with trivial parallelism, as described
in Section~\ref{subsec:number-of-dependencies}.

@@

One advantage of systems with asynchronous execution capabilities is
the ability to mitigate load imbalance, especially in the presence of
task parallelism. To quantify this effect,
Figure~\ref{fig:efficiency-imbalance} plots task granularity vs
efficiency curves under varying amounts of load imbalance. The
imbalanced workloads are generated by randomly varying the task
durations by $1+(X-0.5)I$
where $X$ is a uniform random variable in the range $[0,1]$ and $I$ is
the degree of imbalance in the range $[0,2]$. The random numbers are
generated with a deterministic pseudo random number generator with a
deterministic seed to ensure that the task durations under all systems
are identical.

The MPI implementation of Task Bench, being bulk synchronous,
suffers the most under load imbalance. The biggest
difference is at large task granularities, where the imbalance has the
effect of putting an upper bound on MPI efficiency. At smaller task
granularities the effect shrinks and may even reverse as systems hit
their fundamental limits due to overhead.

The difference between the remaining systems is due primarily to the
scheduling behavior of the various systems. For example, OpenMP
executes tasks via a thread pool, resulting in a form of work
stealing. This on-node work stealing, in combination with the task
parallelism due to the presence of 4 task graphs, is sufficient to
completely mitigate load imbalance. Systems that exploit task
parallelism but not work stealing on a node achieve lower peak
efficiency, but may achieve smaller task granularities as a result of
not relying on a shared global queue for tasks.

@@

Also of interest is the ability of asynchronous systems to
hide communication latency in the presence of task
parallelism. Figure~\ref{fig:efficiency-communication} plots task
granularity vs efficiency curves for varying amounts of
communication. The degree of communication is configured by the number
of bytes produced by each task (and therefore communicated with each
task dependency).
